---
title: "gentrym1_OriginalHomeworkCode_03"
author: "Gentry Miller"
format: html
editor: visual
theme: readthedown
html:
  toc: TRUE
  toc-depth: 2
  toc-title: Contents
---

# Homework 3

## Preliminary

```{r}
#| label: Loading Packages

#install.packages("rmdformats", repos = "http://cran.us.r-project.org")
#library(tidyverse)
#library(rmdformats)
#library(cowplot)
```

```{r}
#| label: Setting up Data Frame

# Read zombies.csv and store data in dataframe df
df <- read.csv("C:/Users/mille/Documents/School/BU Undergrad/Senior/Spring25/AN588/AN588-Homework3-gentrym1/zombies.csv")

df <- df %>% 
  mutate( 
    id = as.numeric(id),
    first_name = as.factor(first_name),
    last_name = as.factor(last_name),
    gender = as.factor(gender),
    height = as.numeric(height),
    weight = as.numeric(weight),
    zombies_killed = as.numeric(zombies_killed),
    years_of_education = as.numeric(years_of_education),
    major = as.factor(major),
    age = as.numeric(age)
  ) %>% 
  rename(education = years_of_education)
```

## Question 1

Calculate the *population* mean and standard deviation for each quantitative random variable (height, weight, age, number of zombies killed, and years of education).

```{r}
#| label: Means and SDs

# defining functions for population variance (pop_var) and
# population standard deviation (pop_sd)
pop_var <- function(x){
  sum((x - mean(x))^2)/(length(x))  # sum of squares over population size
}
pop_sd <- function(x){
  sqrt(pop_var(x))  # sqrt of variance = standard deviation
}

h_mean <- mean(df$height) # mean height
h_sd <- pop_sd(df$height) # sd of height
w_mean <- mean(df$weight) # mean weight
w_sd <- pop_sd(df$weight)
age_mean <- mean(df$age) # mean age
age_sd <- pop_sd(df$age)
kdr_mean <- mean(df$zombies_killed) # mean number of zombies killed
kdr_sd <- pop_sd(df$zombies_killed)
edu_mean <- mean(df$education) # mean number of years of education
edu_sd <- pop_sd(df$education)

# creating matrix of stats for easy access
names <- c("mean height", "height sd", "mean weight", "weight sd", "mean age", "age sd", "mean kills", "kills sd", "mean years of edu", "years of edu sd")
stat_array <- c(h_mean, h_sd, w_mean, w_sd, age_mean, age_sd, kdr_mean, kdr_sd, edu_mean, edu_sd)

pop_stats <- matrix(stat_array, nrow=10, ncol=1, byrow=TRUE)
rownames(pop_stats) <- names
colnames(pop_stats) <- "Value"
pop_stats
```

## Question 2

Use {ggplot} to make boxplots of each of these variables by gender.

```{r}
#| label: Box Plots by Gender

colors = c("plum1","steelblue1", "olivedrab2")

# creating plots to be placed in array
# mean height
p1 <- ggplot(data=df, aes(x = gender, y = height, fill = gender))
p1 <- p1 + geom_boxplot(show.legend = FALSE)
p1 <- p1 + ylab("Height (in)")
p1 <- p1 + xlab("Gender")
p1 <- p1 + scale_fill_manual(values = colors)

# mean weight
p2 <- ggplot(data=df, aes(x = gender, y = weight, fill = gender))
p2 <- p2 + geom_boxplot(show.legend = FALSE)
p2 <- p2 + ylab("Weight (lbs)")
p2 <- p2 + xlab("Gender")
p2 <- p2 + scale_fill_manual(values = colors)

# mean age
p3 <- ggplot(data=df, aes(x = gender, y = age, fill = gender))
p3 <- p3 + geom_boxplot(show.legend = FALSE)
p3 <- p3 + ylab("Age (years)")
p3 <- p3 + xlab("Gender")
p3 <- p3 + scale_fill_manual(values = colors)

# mean zombies killed
p4 <- ggplot(data=df, aes(x = gender, y = zombies_killed, fill = gender))
p4 <- p4 + geom_boxplot(show.legend = FALSE)
p4 <- p4 + ylab("Zombies Killed")
p4 <- p4 + xlab("Gender")
p4 <- p4 + scale_fill_manual(values = colors)

# mean years of education
p5 <- ggplot(data=df, aes(x = gender, y = education, fill = gender))
p5 <- p5 + geom_boxplot(show.legend = FALSE)
p5 <- p5 + ylab("Years of Education")
p5 <- p5 + xlab("Gender")
p5 <- p5 + scale_fill_manual(values = colors)

# Plot grid
pgrid1 <- plot_grid(p1, p2, p3, p4, p5, nrow = 2, ncol = 3)
pgrid1
```

## Question 3

### 3.1

Use {ggplot} to make scatterplots of height and weight in relation to age.

```{r}
#| label: Scatter Plots by Age

# creating plots to go into the grid
p6 <- ggplot(df, aes(x = age, y = height))
p6 <- p6 + geom_point()  # Scatter plot
p6 <- p6 + geom_smooth(method = "lm")  # Line of best fit (linear regression)
p6 <- p6 + ylab("Height (in)")
p6 <- p6 + xlab("Age (years)")

p7 <- ggplot(df, aes(x = age, y = weight))
p7 <- p7 + geom_point()  # Scatter plot
p7 <- p7 + geom_smooth(method = "lm")  # Line of best fit (linear regression)
p7 <- p7 + ylab("Weight (lbs)")
p7 <- p7 + xlab("Age (years)")

# creating plot grid
pgrid2 <- plot_grid(p6, p7)
pgrid2
```

### 3.2

Do these variables seem to be related? In what way?

> Yes, both height and weight appear to be positively correlated with age.

## Question 4

### 4.1

Using histograms and Q-Q plots, check whether the quantitative variables seem to be drawn from a normal distribution.

```{r}
#| label: Histograms and QQ Plots

# height
hist(df$height, breaks = seq(from = 50, to = 85, by = .5), xlab = "Height (in)", probability = TRUE)
qqnorm(df$height, main = "Height QQ Plot")
qqline(df$height, col = "orange")

# weight
hist(df$weight, breaks = seq(from = 90, to = 215, by = 1), xlab = "Weight (lbs)", probability = TRUE)
qqnorm(df$weight, main = "Weight QQ Plot")
qqline(df$weight, col = "orange")

# age
hist(df$age, breaks = seq(from = 10, to = 30, by = 0.5), xlab = "Age", probability = TRUE)
qqnorm(df$age, main = "Age QQ Plot")
qqline(df$age, col = "orange")

# zombie kills
hist(df$zombies_killed, breaks = seq(from = 0, to = 12, by = 1), xlab = "Zombies Killed", probability = TRUE)
qqnorm(df$zombies_killed, main = "Zombie Kills QQ Plot")
qqline(df$zombies_killed, col = "orange")

# years of education
hist(df$education, breaks = seq(from = 0, to = 8, by = 1), xlab = "Years of Education", probability = TRUE)
qqnorm(df$education, main = "Years of Edu. QQ Plot")
qqline(df$education, col = "orange")

# h, w, a, k, e
```

### 4.2

Which seem to be and which do not (hint: not all are drawn from the normal distribution)?

> It seems like Zombies Killed and Years of Education are the only statistics that aren't drawn from the normal distribution.

### 4.3

For those that are not normal, can you determine from which common distribution they are drawn?

> Both are count data, so I think they draw from a Poisson distribution.

## Question 5

### 5.1

Now use the `sample()` function to sample ONE subset of 30 zombie survivors (without replacement) from this population and calculate the mean and sample standard deviation for each variable.

### 5.2

Also estimate the standard error for each variable, and construct the 95% confidence interval for each mean.

Note that for the variables that are not drawn from the normal distribution, you may need to base your estimate of the CIs on slightly different code than for the normal…

## Question 6

### 6.1

Now draw 99 more random samples of 30 zombie apocalypse survivors, and calculate the mean for each variable for each of these samples.

```{r}

```

### 6.2

Together with the first sample you drew, you now have a set of 100 means for each variable (each based on 30 observations), which constitutes a *sampling* distribution for each variable. What are the means and standard deviations of this distribution of means for each variable?

### 6.3

How do the standard deviations of means compare to the standard errors estimated in \[5\]?

### 6.4

What do these sampling distributions look like (a graph might help here)? Are they normally distributed?

What about for those variables that you concluded were *not* originally drawn from a normal distribution?
